---
title: "msca32014_hw4"
author: "Zhiyin Shi"
date: "April 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Binomial output with two groups\
Consider problem of identifying responses of two gender groups (column sex).\
Decide what is the right model and analyze it using JAGS.\
Consider different prior distributions.\

##0. Import related sources and data
```{r}
sourcePath = "/Users/JaneShi/Desktop/MSCA32014/DBDA2EPrograms2"
suppressWarnings(source(paste(sourcePath,"DBDA2E-utilities.R",sep="/")))

dataPath   = "/Users/JaneShi/Desktop/MSCA32014/Lecture4/HW4/"
data       = read.csv(paste(dataPath, "CourseProjectData.csv", sep = "/"))

sex        = as.numeric(data$sex)
thisdata   = data.frame(cbind(y = data[, 6], sex = sex))
y.male     = thisdata[thisdata$sex == 1, 1]
y.female   = thisdata[thisdata$sex == 2, 1]
tot.male   = length(y.male)
tot.female = length(y.female)

dataList.male   = list(y = y.male,   Ntotal = tot.male)
dataList.female = list(y = y.female, Ntotal = tot.female)

(MLE.meanMale   = sum(y.male)/tot.male)
(MLE.meanFemale = sum(y.female)/tot.female)
```

##1. Modeling with different prior distributions for male and female
1.1 Model Description
```{r}
modelString.SingleModel1=" 
model {
  for (i in 1:Ntotal) {
    y[i]~dbern(theta)
  }
  theta~dbeta(2,2)
}
"
writeLines( modelString.SingleModel1 , con="TEMPmodelSingle.txt" )
```

1.2 Initilization function
```{r}
initsList = function(y) {
  resampledY = sample(y, replace=TRUE ) # resample
  thetaInit  = sum(resampledY)/length(resampledY) 
  thetaInit  = 0.001+0.998*thetaInit # keep away from 0,1
  return( list( theta=thetaInit ) )
}

initsList(y.male)
initsList(y.female)
```

1.3 Run MCMC with JAGS for Male Group
```{r}
library(rjags)
jagsModel1 = jags.model(file = "TEMPmodelSingle.txt", data = dataList.male,
                      n.chains= 3, n.adapt= 600, inits= initsList(y.male))

codaSamples = coda.samples(jagsModel1, variable.names = c("theta"), 
                           n.iter = 4000)
```

1.4 Analyze the result for male group
```{r}
head(codaSamples)

#Check summary statistics
summary(codaSamples)

#Check trace and density of theta
plot(codaSamples)

#Check theta autocorrelation
effectiveSize(codaSamples)

#Check sufficiency of convergence through shrinkage factor
gelman.plot(codaSamples)

#Create posterior theta distribution plot
postHistChain1 = hist(unlist(codaSamples[1]), plot = FALSE)
postHistChain2 = hist(unlist(codaSamples[2]), plot = FALSE)
postHistChain3 = hist(unlist(codaSamples[3]), plot = FALSE)

plot(postHistChain1$mids, postHistChain1$density, type= "l", col = "black",
     lwd= 2, ylim =c(0,120), ylab = "Distribution Density", xlab = "Theta")
lines(postHistChain2$mids, postHistChain2$density, type="l", col="red",
      lwd=2)
lines(postHistChain3$mids, postHistChain3$density, type="l", col="blue",
      lwd=2)
lines(postHistChain3$mids, dbeta(postHistChain3$mids, 1 + sum(y.male),
      tot.male- sum(y.male) + 1), type = "l", col = "green", lwd = 3)
legend("topright", legend = c("Chain1","Chain2","Cahin3","Theoretical"),
       col = c("black","red","blue","green"), lwd = 2)

```
**Comment:** The prior distribution in male group is assumed to be beta(2,2). By running MCMC with JAGS, the three chains each with 4000 iterations resulted in mean = 0.3405 and sd = 0.0050. The trace plot as well as gelman plot show great convergence. The density plot of posterior theta distribution shows bell-shape with fat tail. The effective size is 11807.23, indicating nearly no autocorrelation among theta. Comparing theta distribution of each chain with the theoretical distribution, the MCMC gets a very good result.\
\
1.5 Run MCMC with JAGS for Female Group
```{r}
modelString.SingleModel2=" 
model {
  for (i in 1:Ntotal) {
    y[i]~dbern(theta)
  }
  theta~dbeta(3,3)
}
"
writeLines(modelString.SingleModel2, con="TEMPmodelSingle.txt" )

jagsModel2 = jags.model(file= "TEMPmodelSingle.txt", data= dataList.female,
                    n.chains= 3, n.adapt= 600, inits= initsList(y.female))

codaSamples = coda.samples(jagsModel2, variable.names = c("theta"), 
                           n.iter = 4000)
```

1.6 Analyze the result for Female Group
```{r}
head(codaSamples)

#Check summary statistics
summary(codaSamples)

#Check trace and density of theta
plot(codaSamples)

#Check theta autocorrelation
effectiveSize(codaSamples)

#Check sufficiency of convergence through shrinkage factor
gelman.plot(codaSamples)

#Create posterior theta distribution plot
postHistChain1 = hist(unlist(codaSamples[1]), plot = FALSE)
postHistChain2 = hist(unlist(codaSamples[2]), plot = FALSE)
postHistChain3 = hist(unlist(codaSamples[3]), plot = FALSE)

plot(postHistChain1$mids, postHistChain1$density, type= "l", col = "black",
     lwd= 2, ylim =c(0,120), ylab = "Distribution Density", xlab = "Theta")
lines(postHistChain2$mids, postHistChain2$density, type="l", col="red",
      lwd=2)
lines(postHistChain3$mids, postHistChain3$density, type="l", col="blue",
      lwd=2)
lines(postHistChain3$mids, dbeta(postHistChain3$mids, 1 + sum(y.female),
      tot.female- sum(y.female) + 1), type = "l", col = "green", lwd = 3)
legend("topright", legend = c("Chain1","Chain2","Cahin3","Theoretical"),
       col = c("black","red","blue","green"), lwd = 2)
```
**Comment:** The prior distribution in female group is assumed to be beta(3,3). By running MCMC with JAGS, the three chains each with 4000 iterations resulted in mean = 0.3826 and sd = 0.0040. The trace plot as well as gelman plot show great convergence. The density plot of posterior theta distribution shows bell-shape with fat tail. The effective size is 12000, indicating no autocorrelation among theta. Comparing theta distribution of each chain with the theoretical distribution, the MCMC gets a very good result.