---
title: "msca32014_hw6"
author: "Zhiyin Shi"
date: "May 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

0. Import Data
```{r}
suppressWarnings(library(quantmod))
getSymbols("AAPL", from="2015-1-1",to="2015-12-31")
AAPL.1.day = as.matrix(AAPL)
AAPL.returns.2015 = diff(AAPL.1.day[,6])

dataPath = "/Users/JaneShi/Desktop/MSCA32014/Lecture6"
AAPL.ret = read.csv(paste(dataPath,"AAPL_2015.csv",sep="/"))$x
```

1. Visualize the distribution of the returns.
```{r}
AAPL.dens = density(AAPL.ret)
plot(AAPL.dens)
lines(AAPL.dens$x,dnorm(AAPL.dens$x,mean(AAPL.ret),sd(AAPL.ret)),col="red")
```

2. Using the methods of the workshop estimate parameters of normal model and robust normal model for the Apple returns.\
Estimate realized arithmetic volatility of Apple returns using both normal and robust models.\
Volatility is calculated as sigma = Sd * sqrt(365), where Sd is standard deviation of daily arithmetic returns S(t)−S(t−1), S(t) is stock price.\
\
2.1. Estimate parameters of normal model\
2.1.1. Prepare dataList for JAGS model
```{r}
dataList1 = list(
    y       = AAPL.ret ,
    Ntotal  = length(AAPL.ret),
    mean_mu = mean(AAPL.ret) ,
    sd_mu   = sd(AAPL.ret)
)
```

2.1.2 Prepare JAGS Model String
```{r}
suppressWarnings(library(HDInterval))
suppressWarnings(library(rjags))

modelString1 = "
  model {
    for ( i in 1:Ntotal) {
      y[i] ~ dnorm(mu , 1/sigma^2)
    }
    mu ~ dnorm(mean_mu , 1/(100*sd_mu)^2)
    sigma ~ dunif( sd_mu/1000 , sd_mu*1000 )
  }"
# Write out modelString to a text file
writeLines(modelString1 , con= "TEMPmodel.txt")
```

2.1.3. Initialize JAGS
```{r}
initsList = function() {
    upDown = sample(c(1,-1), 1)
    m      = mean(AAPL.ret) * (1 + upDown * .05)
    s      = sd(AAPL.ret) * (1 - upDown * .1) 
    list(mu= m , sigma= s)
}

parameters = c( "mu" , "sigma")     # The parameters to be monitored
adaptSteps = 500                    # Number of steps to "tune" the samplers
burnInSteps= 1000
numSavedSteps=50000
nChains    = 4 
thinSteps  = 1
nIter      = ceiling((numSavedSteps * thinSteps) / nChains)
```

2.1.4 JAGS Kick-off
```{r}
jagsModel1 = jags.model("TEMPmodel.txt", data= dataList1, inits =initsList,
                         n.chains = nChains, n.adapt = adaptSteps)

# Burn-in:
update(jagsModel1 , n.iter=burnInSteps)

# The saved MCMC chain:
codaSamples1 = coda.samples(jagsModel1 , variable.names = parameters, 
                            n.iter = nIter, thin = thinSteps)
```

2.1.5 Analysis
```{r}
summary(codaSamples1)

#See the means
mean(AAPL.ret)
sd(AAPL.ret)

#Convergence plot
plot(codaSamples1)

#Autocorrelation Plot
autocorr.plot(codaSamples1, ask=F)
effectiveSize(codaSamples1)

#Gelman Plot
gelman.diag(codaSamples1)
gelman.plot(codaSamples1)

#HDI Interval
lapply(codaSamples1, function(z) hdi(as.matrix(z)))
```

2.2 Robust estimation using t-distribution\
2.2.1 dataList is the same as dataList1\
2.2.2 Prepare JAGS model String
```{r}
modelString = "
model {
for ( i in 1:Ntotal ) {
    y[i] ~ dt(mu,1/sigma^2,nu)
}
mu ~ dnorm( mean_mu , 1/(100*sd_mu)^2 )
sigma ~ dunif( sd_mu/1000 , sd_mu*1000 )
nu ~ dexp(1/30.0)
}
  " # close quote for modelString
  # Write out modelString to a text file
writeLines( modelString , con="TEMPmodel.txt" )
```

2.2.3 Initialize JAGS
```{r}
initsList2 = function() {
    upDown = sample(c(1,-1), 1)
    m = mean(AAPL.ret) * (1 + upDown * .05)
    s = sd(AAPL.ret) * (1 - upDown * .1) 
    list(mu = m , sigma = s, nu = 2)
}

parameters = c("mu", "sigma", "nu")    # The parameters to be monitored
adaptSteps = 500               # Number of steps to "tune" the samplers
burnInSteps= 1000
nChains    = 3 
thinSteps  = 1
numSavedSteps=50000
(nIter = ceiling( ( numSavedSteps * thinSteps ) / nChains ))
```

2.2.4 JAGS Model
```{r}
jagsModel2 =jags.model("TEMPmodel.txt", data=dataList1, inits=initsList2,
                        n.chains = nChains, n.adapt = adaptSteps)
# Burn-in:
update(jagsModel2, n.iter = burnInSteps)
# The saved MCMC chain:
codaSamples2 = coda.samples(jagsModel2, variable.names = parameters , 
                            n.iter = nIter, thin = thinSteps)
```

2.2.5 Analysis
```{r}
summary(codaSamples2)

#Check stats
mean(AAPL.ret)
sd(AAPL.ret)

plot(codaSamples2)
autocorr.plot(codaSamples2, ask=F)
effectiveSize(codaSamples2)
gelman.diag(codaSamples2)
gelman.plot(codaSamples2)
(HDIofChains<-lapply(codaSamples2,function(z) hdi(as.matrix(z))))
```

2.3 Estimate realized arithmetic volatility of Apple returns using both normal and robust models.
```{r}
first.diff  = diff(AAPL.ret)
n           = length(first.diff)

sd.normal   = sqrt(sum((first.diff - mean(first.diff))^2) / n)
sd.robust   = sd(first.diff)

vol.normal  = sd.normal * sqrt(365)
vol.robust  = sd.robust * sqrt(365)

vol.normal
vol.robust
```

