---
title: "msca32014_hw9"
author: "Zhiyin Shi"
date: "May 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###0. Import Data
```{r}
suppressWarnings(source('/Users/JaneShi/Desktop/MSCA32014/DBDA2Eprograms2/DBDA2E-utilities.R'))  # use function gammaShRaFromModeSD()

anvDat = read.csv("/Users/JaneShi/Desktop/MSCA32014/Lecture9/AnovaShrinkageData.csv")

table(anvDat$Group)

grMeans = aggregate(Y ~ Group, anvDat, mean)
plot(grMeans)

head(anvDat)
```

###1. Model with estimated hyper-parameter
1.1. Fit model described on the diagram.\
Recenter coefficients so that they add up to 0.\
Set up the following contrasts:\
U vs. A\
M vs. A\
G vs. A
```{r}
suppressWarnings(library(rstan))

#Set up dataList
dataList= list(Ntotal = nrow(anvDat), y = anvDat$Y,
               x = as.integer(anvDat$Group), NxLvl = nlevels(anvDat$Group),
               agammaShRa=unlist(gammaShRaFromModeSD(mode=sd(anvDat$Y)/2,
                                                  sd=2*sd(anvDat$Y)))
)

modelString<-"
data {
    int<lower=1> Ntotal;
    real y[Ntotal];
    int<lower=2> NxLvl;
    int<lower=1, upper=NxLvl> x[Ntotal];
    real<lower=0> agammaShRa[2];
}
transformed data {
    real meanY;
    real sdY;
    meanY = mean(y);
    sdY = sd(y);
}
parameters {
    real a0;
    real<lower=0> aSigma;
    vector[NxLvl] a;
    real<lower=0> ySigma;
}
model {
    a0 ~ normal(meanY, 5*sdY);
    aSigma ~ gamma(agammaShRa[1], agammaShRa[2]);
    a ~ normal(0, aSigma);
    ySigma ~ uniform(sdY/100, sdY*10);
    for ( i in 1:Ntotal ) {
        y[i] ~ normal(a0 + a[x[i]], ySigma);
    }
}
generated quantities {
    // Convert a0,a[] to sum-to-zero b0,b[] :
        real b0;
    vector[NxLvl] b;
    b0 = a0 + mean(a);
    b = a - mean(a);
}"

#Set up stan model
stanDso = stan_model( model_code=modelString )

#Set up MCMC model
fit = sampling(stanDso, 
                data=dataList, 
                pars=c('b0', 'b', 'aSigma', 'ySigma'),
                iter=5000, chains = 2, cores = 2)

#Analyze result
library(shinystan)
launch_shinystan(fit)

summary(fit)$summary[,c(1,4,6,8,10)]
cbind(GroupMeans=grMeans,EstimatedMeans=summary(fit)$summary[2:22,1]+summary(fit)$summary[1,1])
plot(fit)
plot(fit,pars=c("b"))
stan_dens(fit)

#Build up contrasts
fit_ext <- rstan::extract(fit)
head(fit_ext$b)
```
1.2. Do any of these contrasts suggest significant differences between the groups?
```{r}
suppressWarnings(library(HDInterval))

#Contrast U vs A
contrast_U_A = fit_ext$b[,1] - fit_ext$b[,21]
plot(contrast_U_A)
hist(contrast_U_A)
(hdiContrast_U_A<-hdi(contrast_U_A))
(sd.contrast_U_A<-sd(contrast_U_A))
(grMean.diff_U_A<-grMeans[grMeans$Group == "U",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,21]))

#Contrast M vs A
contrast_M_A <- fit_ext$b[,1] - fit_ext$b[,13]
plot(contrast_M_A)
hist(contrast_M_A)
(hdiContrast_M_A<-hdi(contrast_M_A))
(sd.contrast_M_A<-sd(contrast_M_A))
(grMean.diff_M_A<-grMeans[grMeans$Group == "M",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,13]))

#Constrast G vs A
contrast_G_A <- fit_ext$b[,1] - fit_ext$b[,7]
plot(contrast_G_A)
hist(contrast_G_A)
(hdiContrast_G_A<-hdi(contrast_G_A))
(sd.contrast_G_A<-sd(contrast_G_A))
(grMean.diff_G_A<-grMeans[grMeans$Group == "G",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,7]))
```
**Comment:** The contrasts suggest significant difference between U & A and M & A. \
\
1.3. For each pair what is the estimated contrast? Compare estimated contrasts with the corresponding group mean differences.
```{r}
est.contrast_U_A<-mean(contrast_U_A)
est.contrast_M_A<-mean(contrast_M_A)
est.contrast_G_A<-mean(contrast_G_A)

res = cbind(c(abs(est.contrast_U_A), abs(est.contrast_M_A), 
              abs(est.contrast_G_A)),
            c(grMean.diff_U_A, grMean.diff_M_A, grMean.diff_G_A))
colnames(res) = c("contrast", "actual")
rownames(res) = c("U_A", "M_A","G_A")
res
```
**Comment:** The estimated contrasts are underestimated, but not much.

###2 Model without hyper-parameter
2.1. Modify the model so that σασα is not estimated from the data. Fit this model
```{r}
dataList= list(Ntotal = nrow(anvDat), y = anvDat$Y,
               x = as.integer(anvDat$Group), NxLvl = nlevels(anvDat$Group))


modelString1<-"
data {
    int<lower=1> Ntotal;
    real y[Ntotal];
    int<lower=2> NxLvl;
    int<lower=1, upper=NxLvl> x[Ntotal];
}
transformed data {
    real meanY;
    real sdY;
    meanY = mean(y);
    sdY = sd(y);
}
parameters {
    real a0;
    vector[NxLvl] a;
    real<lower=0> ySigma;
}
model {
    a0 ~ normal(meanY, 5*sdY);
    a ~ normal(0, 2*sdY);
    ySigma ~ uniform(sdY/100, sdY*10);
    for ( i in 1:Ntotal ) {
        y[i] ~ normal(a0 + a[x[i]], ySigma);
    }
}
generated quantities {
    // Convert a0,a[] to sum-to-zero b0,b[] :
    real b0;
    vector[NxLvl] b;
    b0 = a0 + mean(a);
    b = a - mean(a);
}"

#Set up stan model
stanDso = stan_model( model_code=modelString1)

#Set up MCMC model
fit = sampling(stanDso, 
                data=dataList, 
                pars=c('b0', 'b', 'ySigma'),
                iter=5000, chains = 2, cores = 2)

#Analyze result
library(shinystan)
#launch_shinystan(fit)

summary(fit)$summary[,c(1,4,6,8,10)]
cbind(GroupMeans=grMeans,EstimatedMeans=summary(fit)$summary[2:22,1]+summary(fit)$summary[1,1])
stan_ac(fit, separate_chains = T)
plot(fit)
plot(fit,pars=c("b"))
stan_dens(fit)

#Build up contrasts
fit_ext <- rstan::extract(fit)
head(fit_ext$b)

```
**Comment:** As we can see, the estimation of mean is more accurate to the true values.\
\
2.2. Do any of these contrasts suggest significant differences between the groups?
```{r}
#Contrast U vs A
contrast_U_A = fit_ext$b[,1] - fit_ext$b[,21]
plot(contrast_U_A)
hist(contrast_U_A)
(hdiContrast_U_A<-hdi(contrast_U_A))
(sd.contrast_U_A<-sd(contrast_U_A))
(grMean.diff_U_A<-grMeans[grMeans$Group == "U",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,21]))

#Contrast M vs A
contrast_M_A <- fit_ext$b[,1] - fit_ext$b[,13]
plot(contrast_M_A)
hist(contrast_M_A)
(hdiContrast_M_A<-hdi(contrast_M_A))
(sd.contrast_M_A<-sd(contrast_M_A))
(grMean.diff_M_A<-grMeans[grMeans$Group == "M",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,13]))

#Constrast G vs A
contrast_G_A <- fit_ext$b[,1] - fit_ext$b[,7]
plot(contrast_G_A)
hist(contrast_G_A)
(hdiContrast_G_A<-hdi(contrast_G_A))
(sd.contrast_G_A<-sd(contrast_G_A))
(grMean.diff_G_A<-grMeans[grMeans$Group == "G",2] - grMeans[grMeans$Group == "A",2])
plot(rank(fit_ext$b[,1]),rank(fit_ext$b[,7]))
```
**Comment:** There are significant contrasts among all three pairs.\
\
2.3. For each pair what is the estimated contrast? Compare estimated contrasts with the corresponding group mean differences.
```{r}
est.contrast_U_A<-mean(contrast_U_A)
est.contrast_M_A<-mean(contrast_M_A)
est.contrast_G_A<-mean(contrast_G_A)

res = cbind(c(abs(est.contrast_U_A), abs(est.contrast_M_A), 
              abs(est.contrast_G_A)),
            c(grMean.diff_U_A, grMean.diff_M_A, grMean.diff_G_A))
colnames(res) = c("contrast", "actual")
rownames(res) = c("U_A", "M_A","G_A")
res
```
**Comment:** From the result, we can see much more accurate estimation of contrasts.





```