---
title: "msca32014_hw3"
author: "Zhiyin Shi"
date: "April 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This project helps understanding beta prior distribution for binomial data

#Estimating the Risk of Tumor in a Group of Rats
In example from Section 5.1 the goal was estimation of the probability theta of tumor in a population of female laboratory rats of type “F344” that receive a zero dose of the drug (control group).\
\
In the experiment 4 out of 14 rats developed a tumor.
```{r}
Data<-c(s=14,k=4)
```

Select binomial model yi∼Binom(theta) with probability of tumor theta and beta prior distribution for the parameter theta∼Beta(alpha, beta).\
\
Suppose we know from historical observation of population of “F344” the mean and the variance of beta distribution for theta.\

Using formulas in the interactive demonstration of beta distribution convert mean value miu=0.136 and standard deviation sig=0.1034 of observed empirical probabilities into shapes of beta distribution alpha, beta.\
```{r}
miu = 0.136
sig = 0.1034
commonPart = miu * (1 - miu) / (sig ^ 2) - 1
alpha = miu * commonPart
beta  = (1 - miu) * commonPart
(prior = c(a = alpha, b = beta))
```

Note that this is not a Bayesian approach since it is not based on probability model.\
Calculate parameters of posterior distribution.
```{r}
(posterior<-c(prior[1] + Data[2], Data[1] - Data[2] + prior[2]))
```

Compare prior and posterior distributions and interpret the difference.
```{r}
theta = seq(0,1,length=1000)
matplot(theta, cbind(dbeta(theta,prior[1],prior[2]),
                     dbeta(theta,posterior[1],posterior[2])),
        type = "l", lty = 1, ylab = "prior / posterior distribution")

legend("topright",legend =c("prior","posterior"),lty=1,col=c("blue","red"))

#Compare
(empiricalTheta = Data[2] / Data[1])
```
**Comment:** From experiment, empirical theta = 0.286. The mode of prior distribution of theta is less 0.1 which is biased a lot. The posterior is more than 0.2, which is getting closer to the empirical theta value. So the poterior distribution is moving towards the true distribution.\
\
Assume that probability of tumor in control group is concentrated around unknown mode ω of beta distribution with concentration level of 20.\
ω∼Beta(Aω,Bω)\
Belief based on historical observations is that parameter ω has a tendency to be around 0.4 with concentration 12.\
Use grid approximation to find posterior distribution.\
\
Prior for omega
```{r}
theta   = omega = seq( 0 , 1 , length = 100)
A_omega = 0.4 * (12 - 2) + 1
B_omega = 12 - A_omega

plot(theta, dbeta(theta, A_omega, B_omega))
```

Concentration of theta
```{r}
K = 20
```

Define joint prior function
```{r}
jointPrior<-function(theta, omega, A_omega, B_omega, K){
  res = dbeta(omega, A_omega, B_omega) *
        dbeta(theta, omega * (K-2) + 1, (1 - omega) * (K-2) + 1)
  return(res)
}
```

Calculate joint prior
```{r}
dens = expand.grid(theta, omega)
colnames(dens) = c("theta","omega")

dens$prior = apply(dens, 1, function(z) jointPrior(z[1], z[2], A_omega,
                                                    B_omega, K))
prior.theta.omega = matrix(dens$prior, 100, 100)
prior.theta.omega = prior.theta.omega / sum(prior.theta.omega)
```

Calculate marginals and plot distributions
```{r}
#Mariginal of omega
prior.omega.marginal = apply(prior.theta.omega, 2, sum)
prior.omega.marginal = prior.omega.marginal/sum(prior.omega.marginal) * 100

matplot(omega, cbind(prior.omega.marginal, dbeta(omega, A_omega,
        B_omega)), type = "l", ylab = "Marginal p(omega)")

#Marginal of theta
prior.theta.marginal = apply(prior.theta.omega, 1, sum)
prior.theta.marginal = prior.theta.marginal/sum(prior.theta.marginal) *100 

plot(theta, prior.theta.marginal, type = "l", ylab = "Marginal p(theta)")

#Joint distributoin of theta and omega
persp(theta, omega, prior.theta.omega, d = 1, theta = -25, phi = 20, 
      main="Joint Prior Distribution")

contour(x = omega, y = theta, z = prior.theta.omega, ylab = "omega", 
        xlab = "theta", main="Joint Prior Distribution")
```

Show dependence of theta on omega. Plot 
```{r}
par(mfrow=c(3,1))

#Omega = 0.25
prior.theta.omega.25 = jointPrior(theta, 0.25, A_omega, B_omega, K)
prior.theta.omega.25 = prior.theta.omega.25/sum(prior.theta.omega.25) *100

plot(theta, prior.theta.omega.25, type = "l", ylab = "p(theta|omega=0.25)",
     main = "Marginal prior for Theta")

#Omega = 0.5
prior.theta.omega.5 = jointPrior(theta, 0.5, A_omega, B_omega, K)
prior.theta.omega.5 = prior.theta.omega.5 / sum(prior.theta.omega.5) * 100

plot(theta, prior.theta.omega.5, type = "l", ylab="p(theta|omega=0.5)")

#Omega = 0.75
prior.theta.omega.75 = jointPrior(theta, 0.75, A_omega, B_omega, K)
prior.theta.omega.75 = prior.theta.omega.75 /sum(prior.theta.omega.75) *100

plot(theta, prior.theta.omega.75, type = "l", ylab = "p(theta|omega=0.75)")
```


Calculate likelihood with data of k = 4, s = 14
```{r}
par(mfrow=c(1,1))

likeli = function(theta, s, k){
  theta ^ k * (1 - theta) ^ (s - k)
}

likelihood = likeli(theta, 14, 4)

plot(theta, likelihood, type = "l", ylab = "p(y|theta)", main="Likelihood")
```

Joint posterior distribution and its marginals
```{r}
posterior = apply(prior.theta.omega, 2, function(z) z * likelihood)
posterior = posterior / sum(posterior)

persp(theta, omega, posterior, d = 1, theta = -25, phi = 20, 
      main = "Joint Posterior Distribution")

contour(x = omega, y = theta, z = posterior, ylab = "omega", 
        xlab = "theta", main = "Joint Posterior Distribution")

posterior.omega.marginal = apply(posterior, 2, sum)
posterior.omega.marginal = posterior.omega.marginal /
                           sum(posterior.omega.marginal) * 100

plot(omega, posterior.omega.marginal, type = "l")

posterior.theta.marginal = apply(posterior, 1, sum)
posterior.theta.marginal = posterior.theta.marginal /
                           sum(posterior.theta.marginal) * 100

plot(theta, posterior.theta.marginal, type = "l")
```

Show dependence of theta on omega. 
```{r}
par(mfrow=c(3,1))

#Omega=0.25
post.theta.omega.25 = posterior[, 26]
post.theta.omega.25 = post.theta.omega.25 / sum(post.theta.omega.25) * 100

plot(theta, post.theta.omega.25, type = "l", 
     ylab = "p(theta|omega=0.25,y)", main = "Marginal posterior for Theta")

#Omega=0.5
post.theta.omega.5 = posterior[, 50]
post.theta.omega.5 = post.theta.omega.5 / sum(post.theta.omega.5) * 100

plot(theta, post.theta.omega.5, type = "l", ylab = "p(theta|omega=0.5,y)")

#Omega=0.75
post.theta.omega.75 = posterior[, 75]
post.theta.omega.75 = post.theta.omega.75 / sum(post.theta.omega.75) * 100

plot(theta, post.theta.omega.75, type = "l", ylab ="p(theta|omega=0.75,y)")
```

Compare marginal priors and posteriors for both parameters.
```{r}
par(mfrow=c(1,1))
matplot(theta, cbind(prior.theta.marginal, posterior.theta.marginal),
        type = "l")
matplot(omega, cbind(prior.omega.marginal, posterior.omega.marginal), 
        type = "l")
```