---
title: "msca32014_hw2_zs"
author: "Zhiyin Shi"
date: "April 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This projects helps understanding how Bayes theorem is used in statistical analysis.

###Step1 
Find posterior probability for binomial model with uniform prior and data. Use set.seed(81) for simulation of θ. Find mode, mean and variance of the posterior distribution.\
\
1.1 Simulate data of Bernulli trials
```{r}
set.seed(81)
(data1 = rbinom(10,1,.71))
```

1.2 Construct uniform prior with 1001 theta values
```{r}
theta1   = seq(0.00001, 1 - 0.00001, length = 1001)
p.theta1 = rep(1, length(theta1))
p.theta1 = p.theta1 / sum(p.theta1)
```

1.3 Compute likelihood
```{r}
likeli.binom = function(param, data) {
  num.k = sum(data)
  num.s = length(data)
  likeli.bn = param ^ num.k * (1 - param) ^ (num.s - num.k)
  return(likeli.bn)
}

likelihood1 = likeli.binom(theta1, data1)
```

1.4 Compute posterior distribution
```{r}
post.distr1 = likelihood1 * p.theta1
post.distr1 = post.distr1 / sum(post.distr1)
```

1.5 Plot
```{r}
suppressWarnings(source(paste("/Users/JaneShi/Desktop/MSCA32014/DBDA2Eprograms2/BernGrid.R",sep="/")))

posterior1 = BernGrid(theta1, p.theta1, data1, plotType="Bars", 
                      showCentTend ="None", showHDI =FALSE,
                      showpD =FALSE)
```

1.6 Compute posterior mode, mean and variance
```{r}
mode1 = theta1[ which.max(post.distr1) ]
mean1 = theta1 %*% post.distr1 / sum(post.distr1)
var1  = ((theta1 - mean1) ^ 2) %*% post.distr1 / sum(post.distr1)

c(mode1 = mode1, mean1 = mean1, var1 = var1)
```

###Step2
Add more data and recalculate posterior distribution. Use the set.seed(97) for second simulation of theta. Find mode, mean and variance of the updated posterior distribution.\
\
2.1 Simulate new data
```{r}
set.seed(97)
(data2 = rbinom(10,1,.71))
```

2.2 Update prior
```{r}
p.theta2 = posterior1
```

2.3 Compute likelihood
```{r}
likelihood2 = likeli.binom(theta1, data2)
```

2.4 Compute posterior distribution
```{r}
post.distr2 = likelihood2 * p.theta2
post.distr2 = post.distr2 / sum(post.distr2)
```

2.5 Plot
```{r}
posterior2 = BernGrid(theta1, p.theta2, data2, plotType="Bars", 
                      showCentTend ="None", showHDI =FALSE,
                      showpD =FALSE)
```

2.6 Compute posterior mode, mean and variance
```{r}
mode2 = theta1[ which.max(post.distr2) ]
mean2 = theta1 %*% post.distr2 / sum(post.distr2)
var2  = ((theta1 - mean2) ^ 2) %*% post.distr2 / sum(post.distr2)

c(mode2 = mode2, mean2 = mean2, var2 = var2)
```

###Step3
Repeat steps 1. and 2. using data in reverse order: first use data2, then add data1.\
\
3.1 Define data, prior and likelihood
```{r}
data3       = data2
p.theta3    = p.theta1
likelihood3 = likeli.binom(theta1, data3)
```

3.2 Define posterior distribution
```{r}
post.distr3 = likelihood3 * p.theta3
post.distr3 = post.distr3 / sum(post.distr3)
```

3.3 Plot
```{r}
posterior3 = BernGrid(theta1, p.theta3, data3, plotType="Bars", 
                      showCentTend ="None", showHDI =FALSE,
                      showpD =FALSE)
```

3.4 Compute posterior mode, mean and variance
```{r}
mode3 = theta1[ which.max(post.distr3) ]
mean3 = theta1 %*% post.distr3 / sum(post.distr3)
var3  = ((theta1 - mean3) ^ 2) %*% post.distr3 / sum(post.distr3)

c(mode3 = mode3, mean3 = mean3, var3 = var3)
```

3.5 Update prior with posterior and recalculate posterior. First step, Define data, prior and likelihood.
```{r}
data4       = data1
p.theta4    = posterior3
likelihood4 = likeli.binom(theta1, data4)
```

3.6 Define posterior distribution
```{r}
post.distr4 = likelihood4 * p.theta4
post.distr4 = post.distr4 / sum(post.distr4)
```

3.7 Plot
```{r}
posterior4 = BernGrid(theta1, p.theta4, data4, plotType="Bars", 
                      showCentTend ="None", showHDI =FALSE,
                      showpD =FALSE)
```

3.8 Compute posterior mode, mean and variance
```{r}
mode4 = theta1[ which.max(post.distr4) ]
mean4 = theta1 %*% post.distr4 / sum(post.distr4)
var4  = ((theta1 - mean4) ^ 2) %*% post.distr4 / sum(post.distr4)

c(mode4 = mode4, mean4 = mean4, var4 = var4)
```

###Step4
Compare mode, mean and variance of the updated posterior distributions.
```{r}
post.compare = rbind(c(mode2, mean2, var2), c(mode4, mean4, var4))
rownames(post.compare) = c("data1.then.data2", "data2.then.data1")
colnames(post.compare) = c("mode", "mean", "variance")
post.compare
```
**Comment:** No matter which dataset comes first, the updated posterior distribution is always the same. 
